---
title: "Final Project Stats for Finance"
author: "Raphaël ADDA"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



**Libraries importation**
```{r}
library(tidyverse)
library(tidyquant) # Library which allows to obtain stock prices from yahoo finance
library(tibbletime)
library(ggplot2)
library(moments)
require(reshape2)
library(patchwork)
library(MASS)
library(fGarch)
library(functClust)
library(ggcorrplot)
library(quadprog)
library(Ecdat)
library(ggfortify)
library(bootstrap)
library(copula) 
library(fGarch) 
```




**Creation of a dot plot theme**
```{r}
# create dot plot theme
theme_dotplot <- 
  theme_bw(7) +
  theme(axis.text.y = element_text(size = rel(.8)), axis.ticks.y = element_blank(),
        axis.title.x = element_text(), axis.text = element_text(face = "bold"),
        plot.background = element_rect(fill = "lightcyan2"),
        panel.background = element_rect(fill = "moccasin"),
        panel.grid.major.x = element_line(size = 0.5),
        panel.grid.major.y = element_line(size = 0.5, color = "lightblue"),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(size = rel(.7)), legend.position = "top")
```






## Importation of data

We chose 15 stocks from the 1st of November 2016 to the 1st of December of 2021.
```{r}
tickers = c("AAPL", "NFLX", "AMZN", "JPM","SONY" ,"V","MA","MSFT","ADBE","TSLA","RKUNY","WMT","^DJI","^IXIC","BABA")
daily_prices <- subset(tq_get(tickers,
               from = "2016-09-01",
               to = "2021-12-01",
               get = "stock.prices",period = "monthly"), select = c(date,symbol,close))%>%pivot_wider(names_from = symbol, values_from = close)

tickers = c("Apple", "Netflix", "Amazon", "JP Morgan Chase","Sony" ,"Visa","Mastercard","Microsoft","Adobe","Tesla","Rakuten","Walmart","Dow Jones","Nasdaq","Alibaba")

colnames(daily_prices) = c("date",tickers)
```


We create here the dataframe of the monthly returns, using closing prices of each month.
```{r}
monthly_prices = as_period(as_tbl_time(daily_prices, index = date),'1 month',side = "end")
n = dim(monthly_prices)[1]
returns =  as.data.frame(100 * (as.matrix(monthly_prices[-1])[2:n, ] / monthly_prices[-1][1:(n-1), ] - 1))
mu_f = 0.45
```
mu_f corresponds to the risk free rate.


##Descriptive Statistics:
```{r}
mean_stocks = colMeans(returns)
std_stocks = apply(returns,2,sd)
skewness_stocks = apply(returns,2,skewness)
kurtosis_stocks = apply(returns,2,kurtosis) + 3 #(we add 3 because the kurtosis function 
                                                #computes the normalized kurtosis (by removing 3))

mean_stocks
std_stocks
skewness_stocks
kurtosis_stocks
```


```{r}
df <- melt(monthly_prices ,  id.vars = 'date', variable.name = 'series')

ggplot(df)+
  geom_line(aes(date, value)) + 
  xlab("") + facet_wrap(~series, scales = "free", ncol = 5) + ggtitle("Evolution of monthly prices") + theme_dotplot

# ggsave("Evol_prices.pdf")
```

```{r}
date_col = monthly_prices$date[2:n]
df <- cbind.data.frame(date_col, returns)
df <- melt(df ,  id.vars = 'date_col', variable.name = 'series')
ggplot(df)+
  geom_line(aes(date_col, value)) + 
  xlab("") + facet_wrap(~series, scales = "free", ncol = 5) + ggtitle("Evolution of monthly returns")+ theme_dotplot

# ggsave("Evol_returns.pdf")
```


```{r}
tickers = c("AAPL", "NFLX", "AMZN", "JPM","SONY" ,"V","MA","MSFT","ADBE","TSLA","RKUNY","WMT","^DJI","^IXIC","BABA")
daily_prices <- subset(tq_get(tickers,
               from = "2019-12-01",
               to = "2020-05-01",
               get = "stock.prices",period = "monthly"), select = c(date,symbol,close))%>%pivot_wider(names_from = symbol, values_from = close)

tickers = c("Apple", "Netflix", "Amazon", "JP Morgan Chase","Sony" ,"Visa","Mastercard","Microsoft","Adobe","Tesla","Rakuten","Walmart","Dow Jones","Nasdaq","Alibaba")

colnames(daily_prices) = c("date",tickers)

monthly_prices = as_period(as_tbl_time(daily_prices, index = date),'1 month',side = "end")
n = dim(monthly_prices)[1]
returns =  as.data.frame(100 * (as.matrix(monthly_prices[-1])[2:n, ] / monthly_prices[-1][1:(n-1), ] - 1))
mu_f = 0.45

normalize <- function(x) {
  x/x[1]
}
df <- apply(monthly_prices[tickers] ,2,normalize)
date_col = monthly_prices$date
df <- cbind.data.frame(date_col, df)
df <- melt(df ,  id.vars = 'date_col', variable.name = 'series')
ggplot(df)+
  geom_line(aes(date_col, value)) + 
  xlab("") + facet_wrap(~series, ncol = 5) + ggtitle("Equity curves")+ theme_dotplot

#ggsave("Equity_prices.pdf")
```
Because, we previously modified our dataframe, we recreate it now.
```{r}
tickers = c("AAPL", "NFLX", "AMZN", "JPM","SONY" ,"V","MA","MSFT","ADBE","TSLA","RKUNY","WMT","^DJI","^IXIC","BABA")
daily_prices <- subset(tq_get(tickers,
               from = "2016-09-01",
               to = "2021-12-01",
               get = "stock.prices",period = "monthly"), select = c(date,symbol,close))%>%pivot_wider(names_from = symbol, values_from = close)

tickers = c("Apple", "Netflix", "Amazon", "JP Morgan Chase","Sony" ,"Visa","Mastercard","Microsoft","Adobe","Tesla","Rakuten","Walmart","Dow Jones","Nasdaq","Alibaba")

colnames(daily_prices) = c("date",tickers)

monthly_prices = as_period(as_tbl_time(daily_prices, index = date),'1 month',side = "end")
n = dim(monthly_prices)[1]
returns =  as.data.frame(100 * (as.matrix(monthly_prices[-1])[2:n, ] / monthly_prices[-1][1:(n-1), ] - 1))
mu_f = 0.45
```


```{r}
date_col = monthly_prices$date[2:n]
df <- cbind.data.frame(date_col, returns)
df <- melt(df ,  id.vars = 'date_col', variable.name = 'series')

ggplot(df) + geom_histogram(aes (x =value)) + xlab("") + facet_wrap(~series, scales = "free", ncol = 5) + ggtitle("Histograms of monthly returns")+ theme_dotplot
#ggsave("Histograms.pdf")

```

```{r}
date_col = monthly_prices$date[2:n]
df <- cbind.data.frame(date_col, returns)
df <- melt(df ,  id.vars = 'date_col', variable.name = 'series')

g_median <- ggplot(df) + geom_boxplot(aes (x=value,y=reorder(series,value,median))) + xlab("") + ggtitle("Boxplots of monthly returns, ordered by median")+ theme_dotplot

g_mean <- ggplot(df) + geom_boxplot(aes (x=value,y=reorder(series,value,mean))) + xlab("") + ggtitle("Boxplots of monthly returns, ordered by mean")+ theme_dotplot

g_median + g_mean + plot_layout()
#ggsave("median_boxplots.pdf",g_median)
#ggsave("mean_boxplots.pdf",g_mean)
```



```{r}
date_col = monthly_prices$date[2:n]
df <- cbind.data.frame(date_col, returns)
df <- melt(df ,  id.vars = 'date_col', variable.name = 'series')

ggplot(df,aes(sample =value)) + xlab("") + facet_wrap(~series, ncol = 5) +  
  geom_qq(distribution = stats::qnorm, size = 0.5, col = 'blue', alpha = 0.5)+
  geom_qq_line(distribution = stats::qnorm)+ 
  ggtitle("qq-plots of monthly returns")+ 
  theme_dotplot

#ggsave("qqplots.pdf")
```

```{r}
n = dim(returns)[1]
min_AIC <- function(x){
  std_AIC = 2 * stdFit(x)$objective + 2 * 3
  ged_AIC = 2 * gedFit(x)$objective + 2 * 3
  snorm_AIC = 2 * snormFit(x)$objective + 2 * 3
  min_aic = min(c(std_AIC,ged_AIC,snorm_AIC))
  print(c(std_AIC,ged_AIC,snorm_AIC))
  ind = which.min(min_aic)
  c('std', 'ged', 'snorm')[ind]
}
apply(returns,2,min_AIC)
```

```{r}
n = dim(returns)[1]
min_BIC <- function(x){
  std_BIC = 2 * stdFit(x)$objective + log(n) *  3
  ged_BIC = 2 * gedFit(x)$objective + log(n) *  3
  snorm_BIC = 2 * snormFit(x)$objective + log(n)  * 3
  min_bic = min(c(std_BIC,ged_BIC,snorm_BIC))
  ind = which.min(min_bic)
  c('std', 'ged', 'snorm')[ind]
}
apply(returns,2,min_BIC)
```


Conversion to annual mean and std
```{r}
annual_means = mean_stocks*12
annual_std = std_stocks*sqrt(12)
annual_means
annual_std
```

Pairwise scatter plot

```{r}
pairs(returns)
```

Covariance matrix
```{r}
cov_stocks = cov(returns)
cov_stocks
```

##Portfolio Theory

No shorts allowed, no constraints on quantity
```{r}
#pdf(file="eff_front.pdf")
Amat = cbind(rep(1, length(mean_stocks)), mean_stocks) # set the constraints matrix
Amat = cbind(cbind(Amat,diag(length(mean_stocks))),-diag(length(mean_stocks))) # set the constraints matrix
muP = seq(min(mean_stocks)+0.0001,max(mean_stocks)-0.00001,length = 300) # target portfolio means
# for the expect portfolio return
sdP = muP # set up storage for std dev’s of portfolio returns
weights = matrix(0, nrow = 300, ncol = length(mean_stocks)) # storage for weights
for (i in 1:length(muP)) # find the optimal portfolios
{
  bvec = c(1,muP[i],rep(0,length(mean_stocks)),rep(-1,length(mean_stocks)))# constraint vector
  result = 
    solve.QP(Dmat = 2 * cov_stocks, dvec = rep(0, length(mean_stocks)), Amat = Amat, bvec = bvec, meq = 2)
  
  sdP[i] = sqrt(result$value)
  weights[i,] = result$solution
}
plot(sdP, muP, type = "l", xlim = c(0, 12),ylim = c(0, 6), lty = 3)# plot efficient frontier (and inefficient portfolios below the min var portfolio)
mu_f = 0.45  # input value of risk-free interest rate
points(0, mu_f, cex = 6, pch = ".") # show risk-free asset
sharpe = (muP - mu_f) / sdP # compute Sharpe’s ratios
ind = (sharpe == max(sharpe)) # Find maximum Sharpe’s ratio
lines(c(0, 12), mu_f + c(0, 12) * (muP[ind] - mu_f) / sdP[ind],lwd = 4, lty = 1, col = "blue")
points(sdP[ind], muP[ind], cex = 4, pch = "*")
ind2 = (sdP == min(sdP)) # find minimum variance portfolio
points(sdP[ind2], muP[ind2], cex = 2, pch = "+")
ind3 = (muP > muP[ind2])
lines(sdP[ind3], muP[ind3], type = "l", xlim = c(0, 12),ylim = c(0, 6), lwd = 3, col = "red") # plot efficient frontier
legend(0, 6, legend=c("efficient frontier", "efficient portfolios"),
       col=c("red","black"), lty=c(1,2), cex=0.8)
# + is the mvp portfolio, * is the tangency portfolio
#Efficient frontier (solid, red), line of efficient portfolios (dashed black) connecting the risk-free asset and tangency portfolio (*), and the minimum variance portfolio (+) with all the six stocks. (Theyre are also shown on reward-risk space).

print("The tangency portfolio is (in percents): ")
print(weights[ind, ]*100)
print("The minimum variance portfolio is (in percents): ")
print(weights[ind2, ]*100)
#dev.off()

sd_mvp = sdP[ind2]
mean_mvp = muP[ind2]
sd_t = sdP[ind]
mean_t = muP[ind]
sharp_mvp = (mean_mvp - mu_f)/sd_mvp
sharp_t = (mean_t - mu_f)/sd_t

weights_Tangency = weights[ind, ]*100
mean_mvp
sd_mvp
sharp_mvp
mean_t
sd_t
sharp_t
```

```{r}
sharpe_returns = (mean_stocks - mu_f) / std_stocks
sharpe_returns

sharpe_tangency = max(sharpe)
sharpe_tangency

```


Expected Shortfall and Value at Risk
```{r}
MVP_returns = as.matrix(returns) %*% as.matrix(weights[ind2,]) 
Tangency_returns = as.matrix(returns) %*% as.matrix(weights[ind,]) 
VaR_function <- function(x){
  S0 = 100000
  alpha = 0.05
  q= as.numeric(quantile(x , alpha))
  VaR = -S0 * q/100
  return(VaR)
}

apply(returns,2,VaR_function)
VaR_function(MVP_returns)
VaR_function(Tangency_returns)

ES_function <- function(x){
  S0 = 100000
  alpha = 0.05
  q= as.numeric(quantile(x , alpha))
  IEVaR = (x  < q)
  ES = (-S0 * sum(x  * IEVaR) / sum(IEVaR))/100
  return(ES)
}
apply(returns,2,ES_function)
ES_function(MVP_returns)
```


shorts allowed (limited to -10% per stocks)
```{r}
#pdf(file="eff_front_short.pdf")
Amat = cbind(rep(1, length(mean_stocks)), mean_stocks) # set the constraints matrix
Amat = cbind(cbind(Amat,diag(length(mean_stocks))),-diag(length(mean_stocks))) # set the constraints matrix
muP = seq(min(mean_stocks)+0.0001,max(mean_stocks)-0.00001,length = 300) # target portfolio means
# for the expect portfolio return
sdP = muP # set up storage for std dev’s of portfolio returns
weights = matrix(0, nrow = 300, ncol = length(mean_stocks)) # storage for weights
for (i in 1:length(muP)) # find the optimal portfolios
{
  bvec = c(1,muP[i],rep(-0.1,length(mean_stocks)),rep(-1,length(mean_stocks)))# constraint vector
  result = 
    solve.QP(Dmat = 2 * cov_stocks, dvec = rep(0, length(mean_stocks)), Amat = Amat, bvec = bvec, meq = 2)
  
  sdP[i] = sqrt(result$value)
  weights[i,] = result$solution
}
plot(sdP, muP, type = "l", xlim = c(0, 12),ylim = c(0, 6), lty = 3)# plot efficient frontier (and inefficient portfolios below the min var portfolio)
mu_f = 0.45  # input value of risk-free interest rate
points(0, mu_f, cex = 6, pch = ".") # show risk-free asset
sharpe = (muP - mu_f) / sdP # compute Sharpe’s ratios
ind = (sharpe == max(sharpe)) # Find maximum Sharpe’s ratio
lines(c(0, 12), mu_f + c(0, 12) * (muP[ind] - mu_f) / sdP[ind],lwd = 4, lty = 1, col = "blue")
points(sdP[ind], muP[ind], cex = 4, pch = "*")
ind2 = (sdP == min(sdP)) # find minimum variance portfolio
points(sdP[ind2], muP[ind2], cex = 2, pch = "+")
ind3 = (muP > muP[ind2])
lines(sdP[ind3], muP[ind3], type = "l", xlim = c(0, 12),ylim = c(0, 6), lwd = 3, col = "red") # plot efficient frontier
legend(0, 6, legend=c("efficient frontier", "efficient portfolios"),
       col=c("red","black"), lty=c(1,2), cex=0.8)
# + is the mvp portfolio, * is the tangency portfolio
#Efficient frontier (solid, red), line of efficient portfolios (dashed black) connecting the risk-free asset and tangency portfolio (*), and the minimum variance portfolio (+) with all the six stocks. (Theyre are also shown on reward-risk space).

print("The tangency portfolio is (in percents): ")
print(weights[ind, ]*100)
print("The minimum variance portfolio is (in percents): ")
print(weights[ind2, ]*100)


#dev.off()

sd_mvp_short = sdP[ind2]
mean_mvp_short = muP[ind2]
sd_t_short = sdP[ind]
mean_t_short = muP[ind]
sharp_mvp_short = (mean_mvp - mu_f)/sd_mvp
sharp_t_short = (mean_t - mu_f)/sd_t

mean_mvp_short
sd_mvp_short
sharp_mvp_short
mean_t_short
sd_t_short
sharp_t
```


```{r}
sharpe_returns = (mean_stocks - mu_f) / std_stocks
sharpe_returns

sharpe_tangency = max(sharpe)
sharpe_tangency
```


```{r}
MVP_returns = as.matrix(returns) %*% as.matrix(weights[ind2,]) 
tangency_returns = as.matrix(returns) %*% as.matrix(weights[ind,])
apply(returns,2,VaR_function)
VaR_function(MVP_returns)
VaR_function(Tangency_returns)

apply(returns,2,ES_function)
ES_function(MVP_returns)
```

##Asset Allocation

Without risk free asset
```{r}
mu_p = 0.5
Amat = cbind(rep(1, length(mean_stocks)), mean_stocks) # set the constraints matrix
Amat = cbind(cbind(Amat,diag(length(mean_stocks))),-diag(length(mean_stocks))) # set the constraints matrix
# for the expect portfolio return
weights = rep(0,length(mean_stocks))
bvec = c(1,mu_p,rep(0,length(mean_stocks)),rep(-1,length(mean_stocks))) # constraint vector
result = solve.QP(Dmat = 2 * cov_stocks, dvec = rep(0, length(mean_stocks)), Amat = Amat, bvec = bvec, meq = 2)
std_port = sqrt(result$value)
weights = result$solution

weights*100
std_port
port = as.matrix(returns)%*%weights
VaR_function(port)
ES_function(port)
```

With risk free asset
```{r}
mu_p = 0.5
theta = (mu_p - mu_f)/(mean_t - mu_f)

w = c((1-theta)*100, theta*weights_Tangency)
w
std_port = sd_t*theta
port = cbind(rep(mu_f,n),as.matrix(returns))%*%(w/100)
mean(port)
std_port
sd(port)
VaR_function(port)
ES_function(port)
```


##Principal Component Analysis
Correlation matrix heatmap
```{r}
corr_stocks = cor(returns)
pdf(file="corr_mat.pdf")
ggcorrplot(corr_stocks, hc.order = TRUE, type = "lower",
   lab = TRUE,lab_size = 2)
dev.off()
```
pca
```{r}
pdf(file="pca.pdf")
pca = prcomp(returns, scale = TRUE)
autoplot(pca, data=returns, loadings=TRUE,loadings.label=TRUE)
dev.off()
```

```{r}
pca = prcomp(returns, scale = TRUE)
autoplot(pca, data=returns, loadings=TRUE,loadings.label=TRUE,x=1,y=3)
```

```{r}
pca = prcomp(returns, scale = TRUE)
autoplot(pca, data=returns, loadings=TRUE,loadings.label=TRUE,x=2,y=3)
```

factor analysis
```{r}
factanal(x = returns, factors = 5)
```


##Risk Management

Non-Parametric VaR and ES
```{r}
VaR_function <- function(x){
  S0 = 100000
  alpha = 0.05
  q= as.numeric(quantile(x , alpha))
  VaR = -S0 * q/100
  return(VaR)
}

ES_function <- function(x){
  S0 = 100000
  alpha = 0.05
  q= as.numeric(quantile(x , alpha))
  IEVaR = (x  < q)
  ES = (-S0 * sum(x  * IEVaR) / sum(IEVaR))/100
  return(ES)
}

apply(returns,2,VaR_function)
apply(returns,2,ES_function)
```


Parametric VaR and ES
```{r}
VaR_function_parametric <- function(x){
  S0 = 100000
  alpha = 0.05
  mu = mean(x)
  sigma = sd(x)
  q = qnorm(alpha)
  VaR = -S0 * (mu + q*sigma)/100
  return(VaR)
}

ES_function_parametric <- function(x){
  S0 = 100000
  alpha = 0.05
  q = qnorm(alpha)
  ES = S0*(-mu + sigma * dnorm(q)/alpha)/100
  return(ES)
}

apply(returns,2,VaR_function_parametric)
apply(returns,2,ES_function)
```


##Copulas
```{r}
ranker <- function(x){
  rank(x)/(length(x)+1)
}
df = apply(returns,2,ranker)
m = length(tickers)
fnorm = fitCopula(copula=normalCopula(dim=m),data=df,method="ml")
ffrank = fitCopula(copula = frankCopula(3, dim = m),
                     data = df, method = "ml")
fclayton = fitCopula(copula = claytonCopula(1, dim=m),
                     data = df, method = "ml")
fgumbel = fitCopula(copula = gumbelCopula(3, dim=m),
                    data = df, method = "ml")
fjoe = fitCopula(copula=joeCopula(2,dim=m),data=df,method="ml")

fnorm
ffrank
fclayton
fgumbel
fjoe

```
